--- 
output:
  pdf_document:
    keep_tex: true
    fig_caption: true
    toc: true
    toc_depth: 3 
    number_sections: true
    citation_package: natbib
    latex_engine: pdflatex
    template: ./../latex-templates/report.tex
  html_document:
    df_print: paged
    
title: "Will Smith VS Denzel Washington"
subtitle: "Which Actor is better?"
author: 
- name: "Nic Trout"
  affiliation: "Washington State University"
keywords: |
    Boxplot, Correlation Table, T tests
abstract: |
  \noindent To describe who is a better actor, Will or Denzel?
sectionnumberdepth: 3
titleEndnotes: "ENDNOTES"
titleReferences: "REFERENCES"
columnsReferences: 2
titleTOC:  "TABLE OF CONTENTS"
bibliography: ./../biblio/master.bib
bibliostyle: ./../biblio/ormsv080.bst
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section{Introduction}
\label{sec:intro}




\begin{figure}[!ht]
	\hrule
	\begin{center}
	    \scalebox{1.00}{	\includegraphics[trim = 0 0 0 0,clip,width=0.85\textwidth]{pdfs/oneimage.pdf} }
	\end{center}
	\label{fig:oneimage-1}
	\caption{ \textbf{One Image Description:} On the left represents a NBA rookies hand width vs height proportions of hand. On the right you can see your average Joe (the student volunteers measured).}
	\hrule
\end{figure}



\section{Research Question:  What is my primary question}
\label{sec:rq}
Who is better, Will Smith or Denzel Washington?


\section{Data Description}
\label{sec:data}




\subsection{Summary of Sample}
\label{sec:data-sample}




\subsection{Summary Statistics of Data}
\label{sec:data-summary}

 

\section{Key Findings}
\label{sec:findings}



\section{Conclusion}
\label{sec:conclusion}



\newpage



```{r, setup, include=FALSE}

library(devtools);       # required for source_url
library(humanVerseWSU);

path.humanVerseWSU = "https://raw.githubusercontent.com/MonteShaffer/humanVerseWSU/"

source_url( paste0(path.humanVerseWSU,"master/misc/functions-project-measure.R") );

source_url = paste0(path.humanVerseWSU, "master/misc/functions-nlp.R");

source_url = paste0(path.humanVerseWSU, "master/misc/functions-nlp-str.R");

source_url = paste0(path.humanVerseWSU, "master/misc/functions-nlp-stack.R");

source_url = paste0(path.humanVerseWSU, "master/misc/functions-nlp-pos.R");

source_url = paste0(path.humanVerseWSU,"humanVerseWSU/R/functions-encryption.R");

path.to.final = "C:/Users/Nic Trout/Documents/C/WSU_STATS419_FALL2020/final/";

```

```{r, data, include=FALSE}
library(imdb);

imdb::loadDataIMDB();
names(imdb.data);

humanVerseWSU::loadInflationData();

path.to.nascent = "C:/Users/Nic Trout/Documents/C/WSU_STATS419_FALL2020/nascent/";

top2000 = readRDS(paste0(path.to.nascent, 'movieRank2000.rds'));

```


```{r}

top2000$budget = as.numeric(top2000$budget)/1000000;

top2000 = standardizeDollarsInDataFrame(top2000, 2000, "budget", "year", "budget2000");

top2000$budget2000s = 100*(top2000$budget2000/max(top2000$budget2000, na.rm = TRUE));

budget = removeAllColumnsBut(top2000,c("year","budget", "budget2000", "budget2000s", "ttid", "title"));

```

```{r}


top2000$world.gross = as.numeric(top2000$world.gross)/1000000;

top2000 = standardizeDollarsInDataFrame(top2000, 2000, "world.gross", "year", "world.gross2000");

top2000$world.gross2000s = 100*(top2000$world.gross2000/max(top2000$world.gross2000, na.rm = TRUE));

world.gross = removeAllColumnsBut(top2000,c("year","world.gross", "world.gross2000", "world.gross2000s", "ttid", "title"));


my.df = merge(budget,world.gross,by="ttid");

my.df$profit = my.df$world.gross2000 - my.df$budget2000;

my.df$ROI = my.df$profit/my.df$budget2000;


```

```{r}
will.budget = merge(will.movies[,c(1,5)], budget, by = "ttid"); #19/111
will.median = median(will.budget$budget2000, na.rm = TRUE);
will.top = will.budget$ttid
will.ROI = merge(will.movies[,c(1,5)], my.df, by = "ttid"); #19/111


denzel.budget = merge(denzel.movies[,c(1,5)], budget, by = "ttid"); #18/61
denzel.median = median(denzel.budget$budget2000, na.rm = TRUE);
denzel.top = denzel.budget$ttid
denzel.ROI = merge(denzel.movies[,c(1,5)], my.df, by = "ttid"); #18/61


boxplot(will.budget$budget2000, denzel.budget$budget2000);
boxplot(will.budget$budget2000s, denzel.budget$budget2000s);

t.test(will.budget$budget2000,denzel.budget$budget2000);

boxplot(will.ROI$ROI, denzel.ROI$ROI,main = "Will vs Denzel ROI", ylab = "ROI");

#t.test(will.ROI$ROI, denzel.ROI$ROI);


par(mfrow =c(1,2));

boxplot(will.budget$budget2000, main = "Will vs Denzel Budgets", ylim = c(0, 175), ylab = "budget", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.budget$budget2000, main = "", ylim = c(0, 175), ylab = "", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));

par(mfrow =c(1,2));

boxplot(will.ROI$ROI, main = "Will vs Denzel ROI", ylim = c(0, 10), ylab = "", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.ROI$ROI, main = "", ylim = c(0, 10), ylab = "ROI", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));


```



### Rankings: Lots of Ways to Internally Rank
```{r, data-clean-up}

will.search = IMDB.searchPersonName("Will* Smith*");

will.search;

will.nmid = will.search$nmid[1];
will.nmid = "nm0000226";
will.movies = IMDB.getMoviesForPerson(will.nmid);
will.movies;
will.movies = standardizeDollarsInDataFrame(will.movies, 2000, "millions", "year", "millions2000");
will.movies.ttid = will.movies$ttid;


will.not.top = setdiff(will.movies.ttid, will.top);
will.movies$top = "NOT (92)";
#will.movies$top[will.movies$ttid==will.top];
comparison.will = c();
for(i in 1:111)
{
  if(is.element(will.movies$ttid[i], will.top))
  {
    comparison.will = c(comparison.will,i)
  }
}
will.movies$top[comparison.will] = "Top 50 of year (n=19)";

boxplot(millions2000~top,data = will.movies, main = "Income of top vs non-top movies Will", ylim = c(0, 340), frame = F);


```

```{r}



denzel.search = IMDB.searchPersonName("Denzel* Washington*");

denzel.nmid = denzel.search$nmid[1];
denzel.nmid = "nm0000243";
denzel.movies = IMDB.getMoviesForPerson(denzel.nmid);
denzel.movies;
denzel.movies.ttid = denzel.movies$ttid;


denzel.movies = standardizeDollarsInDataFrame(denzel.movies, 2000, "millions", "year", "millions2000");


will.movies = standardizeDollarsInDataFrame(will.movies, 2000, "millions", "year", "millions2000");


denzel.movies.p50 = intersect(denzel.movies.ttid, imdb.data$movies$popular50$ttid);


will.movies.p50 = intersect(will.movies.ttid, imdb.data$movies$popular50$ttid);

denzel.not.top = setdiff(denzel.movies.ttid, denzel.top);
denzel.movies$top = "NOT (n=43)";
#denzel.movies$top[denzel.movies$ttid==denzel.top];
comparison.denzel = c();
for(i in 1:61)
{
  if(is.element(denzel.movies$ttid[i], denzel.top))
  {
    comparison.denzel = c(comparison.denzel,i)
  }
}
denzel.movies$top[comparison.denzel] = "Top 50 of year (n=18)";

par(mfrow = c(1,2));

boxplot(millions2000~top,data = will.movies, main = "Will", ylim = c(0, 340), frame = F);
boxplot(millions2000~top,data = denzel.movies, ylim = c(0, 340), main = "Denzel", frame = F);

par(mfrow = c(1,1));

denzel.p50 = subsetDataFrame(denzel.movies, mycols = "votes" , comparison = ">", myvals = 110000);

will.p50 = subsetDataFrame(will.movies, mycols = "votes" , comparison = ">", myvals = 170000);


denzel.movies;
will.movies;

denzel.movies.p50;
will.movies.p50;
#this shows will killing it
t.test(will.movies$millions2000[comparison.will], denzel.movies$millions2000[comparison.denzel]);
#this shows they are equal on bad movies
t.test(will.movies$millions2000[-comparison.will], denzel.movies$millions2000[-comparison.denzel]);

```

```{r, money-plots}

par(mfrow =c(1,2));

boxplot(will.movies$millions2000, main = "Will vs Denzel Money", ylim = c(0, 375), ylab = "", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.movies$millions2000, main = "", ylim = c(0, 375), ylab = "Millions(2000)", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));

par(mfrow =c(1,2));

boxplot(will.p50$millions2000, main = "Will vs Denzel Popular 50", ylim = c(0, 375), ylab = "", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.p50$millions2000, main = "", ylim = c(0, 375), ylab = "Millions(2000)", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));

plot(denzel.movies$year, denzel.movies$millions2000, pch = 10, cex = 0.75, main = "Denzel's Money in Millions Adjusted(2000)", xlab = "Year", ylab = "Millions", bty = "n", xlim = c(1980,2019), ylim = c(0, 375));
reg.m = lm(denzel.movies$millions2000~denzel.movies$year);
abline(reg.m, col= "blue");

plot(will.movies$year, will.movies$millions2000, pch = 10, cex = 0.75, main = "Will's Money in Millions Adjusted(2000)", xlab = "Year", ylab = "Millions", bty = "n", xlim = c(1980,2019), ylim = c(0, 375));
reg.m = lm(will.movies$millions2000~will.movies$year);
abline(reg.m, col= "blue");

plot(denzel.p50$year, denzel.p50$millions2000, pch = 10, cex = 0.75, main = "Denzel's Popular 50 Money in Millions Adjusted(2000)", xlab = "Year", ylab = "Millions", bty = "n", xlim = c(1980,2019), ylim = c(0, 375));
reg.m = lm(denzel.p50$millions2000~denzel.p50$year);
abline(reg.m, col= "blue");

plot(will.p50$year, will.p50$millions2000, pch = 10, cex = 0.75, main = "Will's Popular 50 Money in Millions Adjusted(2000)", xlab = "Year", ylab = "Millions", bty = "n", xlim = c(1980,2019), ylim = c(0, 375));
reg.m = lm(will.p50$millions2000~will.p50$year);
abline(reg.m, col= "blue");

```


```{r, ratings-plots}
par(mfrow =c(1,2));

boxplot(will.movies$ratings, main = "Will vs Denzel Ratings", ylim = c(0, 10), ylab = "ratings", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.movies$ratings, main = "", ylim = c(0, 10), ylab = "ratings", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));
par(mfrow = c(1,1));

par(mfrow =c(1,2));

boxplot(will.p50$ratings, main = "Will vs Denzel Popular 50 Ratings", ylim = c(0, 10), ylab = "ratings", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.p50$ratings, main = "", ylim = c(0, 10), ylab = "ratings", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));

plot(denzel.movies$year, denzel.movies$ratings, pch = 10, cex = 0.75, main = "Denzel's Ratings", xlab = "Year", ylab = "ratings", bty = "n", xlim = c(1980,2019), ylim = c(0, 10));
reg.m = lm(denzel.movies$ratings~denzel.movies$year);
abline(reg.m, col= "blue");

plot(will.movies$year, will.movies$ratings, pch = 10, cex = 0.75, main = "Will's Ratings", xlab = "Year", ylab = "ratings", bty = "n", xlim = c(1980,2019), ylim = c(0,10));
reg.m = lm(will.movies$ratings~will.movies$year);
abline(reg.m, col= "blue");

plot(denzel.p50$year, denzel.p50$ratings, pch = 10, cex = 0.75, main = "Denzel's Popular 50 Ratings", xlab = "Year", ylab = "ratings", bty = "n", xlim = c(1980,2019), ylim = c(0, 10));
reg.m = lm(denzel.p50$ratings~denzel.p50$year);
abline(reg.m, col= "blue");

plot(will.p50$year, will.p50$ratings, pch = 10, cex = 0.75, main = "Will's Popular 50 Ratings", xlab = "Year", ylab = "ratings", bty = "n", xlim = c(1980,2019), ylim = c(0, 10));
reg.m = lm(will.p50$ratings~will.p50$year);
abline(reg.m, col= "blue");

```



```{r, metacritic-plots}
par(mfrow =c(1,2));

boxplot(will.movies$metacritic, main = "Will vs Denzel Metacritic", ylim = c(0, 100), ylab = "metacritic", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.movies$metacritic, main = "", ylim = c(0, 100), ylab = "metacritic", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));
par(mfrow = c(1,1));

par(mfrow =c(1,2));

boxplot(will.p50$metacritic, main = "Will vs Denzel Popular 50 metacritic", ylim = c(0, 100), ylab = "", xlab = "Will", frame.plot = FALSE);
boxplot(denzel.p50$metacritic, main = "", ylim = c(0, 100), ylab = "metacritic", xlab = "Denzel", frame.plot = FALSE);

par(mfrow = c(1,1));

plot(denzel.movies$year, denzel.movies$metacritic, pch = 10, cex = 0.75, main = "Denzel's metacritic", xlab = "Year", ylab = "metacritic", bty = "n", xlim = c(1980,2019), ylim = c(0, 100));
reg.m = lm(denzel.movies$metacritic~denzel.movies$year);
abline(reg.m, col= "blue");

plot(will.movies$year, will.movies$metacritic, pch = 10, cex = 0.75, main = "Will's metacritic", xlab = "Year", ylab = "metacritic", bty = "n", xlim = c(1980,2019), ylim = c(0, 100));
reg.m = lm(will.movies$metacritic~will.movies$year);
abline(reg.m, col= "blue");

plot(denzel.p50$year, denzel.p50$metacritic, pch = 10, cex = 0.75, main = "Denzel's Popular 50 metacritic", xlab = "Year", ylab = "metacritic", bty = "n", xlim = c(1980,2019), ylim = c(0, 100));
reg.m = lm(denzel.p50$metacritic~denzel.p50$year);
abline(reg.m, col= "blue");

plot(will.p50$year, will.p50$metacritic, pch = 10, cex = 0.75, main = "Will's Popular 50 metacritic", xlab = "Year", ylab = "metacritic", bty = "n", xlim = c(1980,2019), ylim = c(0, 100));
reg.m = lm(will.p50$metacritic~will.p50$year);
abline(reg.m, col= "blue");

```

```{r, some-info-i-pulled}
summary(denzel.movies$millions2000);
summary(will.movies$millions2000);

summary(denzel.movies$ratings);
summary(will.movies$ratings);

summary(denzel.movies$metacritic);
summary(will.movies$metacritic);

#will info
length( intersect(will.movies.ttid, imdb.data$movies$all) );
length( intersect(will.movies.ttid, imdb.data$movies$top250) );
length( intersect(will.movies.ttid, imdb.data$movies$popular50$ttid) );
length( intersect(will.movies.ttid, imdb.data$movies$gem50$ttid) );

#denzel info
length( intersect(denzel.movies.ttid, imdb.data$movies$all) );
length( intersect(denzel.movies.ttid, imdb.data$movies$top250) );
length( intersect(denzel.movies.ttid, imdb.data$movies$popular50$ttid) );
length( intersect(denzel.movies.ttid, imdb.data$movies$gem50$ttid) );
```

