---
title: 'R Notebook: natural language processing (text as multivariate data)'
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
    number_sections: true 
---

```{r}

library(devtools);

library(humanVerseWSU);

path.github = "https://raw.githubusercontent.com/MonteShaffer/humanVerseWSU/master/";

include.me = paste0(path.github, "misc/functions-nlp.R");
# source_url( include.me );
include.me = paste0(path.github, "humanVerseWSU/R/functions-encryption.R");
source_url( include.me );



path.to.nascent = "C:/Users/Alexander Nevsky/Dropbox/WSU-419/Fall 2020/__student_access__/unit_02_confirmatory_data_analysis/nascent/";

folder.nlp = "nlp/";
path.to.nlp = paste0(path.to.nascent, folder.nlp);


```

# Natural Language Processing

There was quite a bit of interest on this topic.  How can we perform multivariate analysis on text.  Most think of data as numbers, but text is certainly data and certainly multivariate.

Below is an email I recently sent to a student not in our program which stands as a good summary of the topic.

<pre style="white-space: pre-wrap; white-space: -moz-pre-wrap; white-space: -pre-wrap; white-space: -o-pre-wrap; word-wrap: break-word;">
Hi M****,

I am in the process of building a few "NLP" notebooks for my class.  I can share those with you in a few weeks.

I have used python in the past for a large project, but it seems like R has more than caught up.

I am leaning towards the package quanteda, but there are several libraries out there:

https://quanteda.io/articles/pkgdown/comparison.html

I like this library because it looks like it is "most hackable" - I have developed a custom NLP-LSA approach and believe I can plug it into this library.

This link compares two books "Sense and Sensibility" and "Moby Dick" ....   A good walk through on some basics of NLP.

https://quanteda.io/articles/pkgdown/replication/digital-humanities.html

I think those two links should help you get started.
</pre>

## Choosing "quanteda"

I have reviewed and tested 3 packages available in R.  I will want to do my own thing often, and so the package needs to be "hackable".  For example, I may want to add my own collection of stop words.  Or I may want to write my own "n-gram" text processing algorithm.  Such elements are things I performed on a "concept-search engine" using patent data to answer the question:  "is your idea patentable?"  This topic will be an (optional) mastery notebook coming soon.



```{r}
## Remember that RStudio links to some files and makes "installation" a bit more difficult than it needs to be.  I would recommend closing RStudio and installing the packages in RGui, then come back to RStudio to use them.


library(devtools);

library(quanteda);  # install.packages("quanteda", dependencies=TRUE); 
library(quanteda.corpora); # devtools::install_github("quanteda/quanteda.corpora");
library(quanteda.textmodels); # install.packages("quanteda.textmodels", dependencies=TRUE); 
library(LIWCalike); # devtools::install_github("kbenoit/LIWCalike");


```


